---
output:
  md_document:
    variant: commonmark
---

```{r include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  error = FALSE
)
```

Steps to get started with dplyr & Postgres, using dbplyr.

## Postgres Setup

[Don’t install Postgres. Docker pull Postgres](https://hackernoon.com/dont-install-postgres-docker-pull-postgres-bee20e200198)

To avoid installing Postgres locally use Docker.

Pull image for the latest stable release:

```{bash eval=FALSE}
docker pull postgres
```

Start Postgres:

```{bash eval=FALSE}
docker run \
  -d --rm \
  --name pg-docker \
  -e POSTGRES_PASSWORD=docker \
  -p 127.0.0.1:5432:5432 \
  postgres
```

## db Connection

### Setup

The [DBI package](https://github.com/r-dbi/DBI) provides a simple interface for connecting to databases and, along with the [odbc package](https://github.com/r-dbi/odbc), makes it easy to set up connections to any with [Open Database Connectivity (ODBC)](https://docs.microsoft.com/en-us/sql/odbc/microsoft-open-database-connectivity-odbc?redirectedfrom=MSDN&view=sql-server-ver15) drivers.

Install odbc dependencies and Postgres driver:

```{bash eval=FALSE}
sudo apt-get install unixodbc unixodbc-dev
sudo apt-get install odbc-postgresql
```

Install DBI and odbc:

```{r eval=FALSE}
install.packages(c('DBI','odbc'))
```

Check odbc package can find Postgres driver:

```{r}
# list unique driver names
unique(odbc::odbcListDrivers()['name'])
```

### Connection

Connect to Postgres database:

```{r}
# load DBI package
library(DBI)

# create connection
con <- dbConnect(
  odbc::odbc(),
  Driver = 'PostgreSQL Unicode',
  Server = '127.0.0.1',
  Database = 'postgres',
  UID = 'postgres',
  PWD = 'docker',
  Port = 5432
)
```

## dbplyr

[Introduction to dbplyr](https://dbplyr.tidyverse.org/articles/dbplyr.html)

Install tidyverse packages (and nycflights13 data):

```{r eval=FALSE}
install.packages(c('tidyverse', 'nycflights13'))
```

Use dbplyr to load nycflights13 data to database:

```{r}
# load tidyverse packages
library(tidyverse)

# copy data to database
con %>%
  copy_to(nycflights13::flights, 'flights')
```

dbplyr interacts with database tables through table references:

```{r}
# create table reference to 'flights' table
flights <- tbl(con, 'flights')

# return top of flights table
head(flights)
```

Once a table reference has been created you can use dplyr to interrogate the table.

dplyr tries to interact with the database as little as possible:

```{r}
# this will not touch the database
flights_jfk <-
  flights %>%
  filter(origin == 'JFK')

# this will
flights_jfk %>%
  ggplot() +
  geom_boxplot(aes(carrier, distance))
```

```{bash include=FALSE}
# clean up
docker stop pg-docker
```
